An Accurate Model for SARS-CoV-2 Pooled RT-PCR Test Errors

### 2021-09-09 paper.tex

\subsection*{Our model}\label{subsec:ours}
Since the essence of the refuted common assumption is that
amplification of all samples occurs only once, we assume the opposite:
amplification of viral RNA succeeds or fails for each sample
independently. Furthermore, according to \cite{Simplistic1,
  Simplistic2, Kim, OptimalDorfmanPool}, a false-positive does not
depend on the number of negative samples in a pool either. For lack of
data pointing otherwise, we incorporate this assumption into our model
with a small modification. We do assume that a false amplification can
occur only once per pool. However, we also assume false amplification
is independent of any other correct amplification. Specifically, it is
possible that every correct amplification fails \emph{and} an
erroneous one occurs simultaneously. This assumption is somewhat
specific for the current application of screening for SARS-CoV-2 via
RT-PCR. For example, cross-reactivity with other coronaviruses would
have violated this assumption. However, cross-reactivity was ruled out
in \cite{KitComparison}. These assumptions lead to our model, which is
illustrated in Figure~\ref{fig:likelihood} and summarized in
\eqref{eq:likelihood}.
\begin{align}\label{eq:likelihood}
    \Pr(\data=0 | \theta) &= \spec\prod_{i=1}^N (1-\sens)^{\theta_i} =
    \spec(1-\Se)^{\sum_i\theta_i}.
\end{align}

@@

To conclude, pooling is an important technique that can improve
testing throughput in a cost-effective manner. Nevertheless, care must
be given to pooling schemes' false-negative rates, especially under
low infection prevalence settings.

>>>>

\subsection*{Our model}\label{subsec:ours}
Since the essence of the refuted common assumption is that
amplification of all samples occurs only once, we assume a more
realistic model: amplification of viral RNA succeeds or fails for each
sample independently. Furthermore, according to \cite{Simplistic1,
  Simplistic2, Kim, OptimalDorfmanPool}, a false-positive does not
depend on the number of negative samples in a pool either. For lack of
data pointing otherwise, we incorporate this assumption into our model
with a small modification. We do assume that a false amplification can
occur only once per pool. However, we also assume false amplification
is independent of any other correct amplification. Specifically, it is
possible that every correct amplification fails \emph{and} an
erroneous one occurs simultaneously. This assumption is somewhat
specific for the current application of screening for SARS-CoV-2 via
RT-PCR. For example, cross-reactivity with other coronaviruses would
have violated this assumption. However, cross-reactivity was ruled out
in \cite{KitComparison}. These assumptions lead to our model, which is
illustrated in Figure~\ref{fig:likelihood} and summarized in
\eqref{eq:likelihood}.
\begin{align}\label{eq:likelihood}
    \Pr(\data=0 | \theta) &= \spec\prod_{i=1}^N (1-\sens)^{\theta_i} =
    \spec(1-\Se)^{\sum_i\theta_i}.
\end{align}

@@

To conclude, pooling is an important technique that can increase
testing throughput in a cost-effective manner. Nevertheless, care must
be given to pooling schemes' false-negative rates, especially under
low infection prevalence settings.
